---
title: "Scorecard Feedback Dashboard"
subtitle: "Open-text survey analysis"
format:
  dashboard:
    theme: cosmo
    css: styles.css
    orientation: rows
    scrolling: true
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
library(tidyverse)
library(readxl)
library(tidytext)
library(plotly)
library(DT)
library(scales)

theme_set(theme_minimal(base_size = 13) +
  theme(panel.grid.minor = element_blank(), plot.title = element_text(face = "bold")))

teal <- "#0F766E"
navy <- "#183153"
amber <- "#D97706"
red <- "#B91C1C"
gray <- "#64748B"

file_path <- "data/synthetic_scorecard_survey_responses.xlsx"
survey <- read_excel(file_path, sheet = "Scorecard Responses")

question_map <- c(
  overall = "What were your overall thoughts on the scorecard?",
  liked = "What, if anything, did you like most about it?",
  unclear = "What, if anything, was problematic or unclear?",
  results = "What did you think of the output/results of the scorecard?",
  guide = "Do you have any feedback on the step-by-step guide?",
  other = "Is there any other feedback you would like to provide after using the scorecard?"
)

survey_long <- survey |>
  select(respondent_id, department, all_of(unname(question_map))) |>
  pivot_longer(all_of(unname(question_map)), names_to = "question_text", values_to = "response") |>
  mutate(
    question = names(question_map)[match(question_text, unname(question_map))],
    question = factor(question, levels = names(question_map)),
    response = str_squish(response)
  ) |>
  filter(!is.na(response), response != "")

theme_rules <- tribble(
  ~theme, ~pattern,
  "Clear and useful", "\\bclear\\b|useful|\\beasy\\b|helpful|practical|straightforward|structured",
  "Scoring needs explanation", "\\bscore\\b|\\bscoring\\b|criteria|rating|borderline|calculation",
  "Too long or repetitive", "long|time|repetitive|repeated|shorter|too many",
  "Technical language", "technical|terminology|jargon|plain.language|glossary|provenance|metadata",
  "Needs recommendations", "recommend|next step|action|remediation|resource|what action",
  "Navigation and saving", "navigation|click|save|progress|screen|move between",
  "Collaboration and roles", "team|role|collaborat|data owner|project manager",
  "Result validity", "confiden|valid|missing answer|not.applicable|completeness|uncertainty"
)

theme_hits <- survey_long |>
  crossing(theme_rules) |>
  mutate(hit = str_detect(str_to_lower(response), pattern)) |>
  filter(hit) |>
  distinct(respondent_id, department, question, response, theme)

bing <- get_sentiments("bing")
sentiment <- survey_long |>
  unnest_tokens(word, response) |>
  inner_join(bing, by = "word") |>
  count(respondent_id, department, question, sentiment) |>
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) |>
  mutate(
    score = positive - negative,
    sentiment_group = case_when(score > 0 ~ "Positive", score < 0 ~ "Negative", TRUE ~ "Mixed/neutral")
  )

sentiment_complete <- survey_long |>
  distinct(respondent_id, department, question) |>
  left_join(sentiment, by = c("respondent_id", "department", "question")) |>
  mutate(sentiment_group = replace_na(sentiment_group, "Mixed/neutral"))

words <- survey_long |>
  unnest_tokens(word, response) |>
  anti_join(stop_words, by = "word") |>
  filter(str_length(word) > 2, !word %in% c("scorecard", "dataset", "guide", "result", "results"))

bigrams <- survey_long |>
  unnest_tokens(bigram, response, token = "ngrams", n = 2) |>
  separate(bigram, c("word1", "word2"), sep = " ", remove = FALSE) |>
  filter(!word1 %in% stop_words$word, !word2 %in% stop_words$word) |>
  count(question, bigram, sort = TRUE)

completion_rate <- survey_long |>
  distinct(respondent_id, question) |>
  nrow() / (n_distinct(survey$respondent_id) * length(question_map))

plotly_bar <- function(data, x, y, title, color = teal) {
  ggplotly(
    ggplot(data, aes(x = {{x}}, y = reorder({{y}}, {{x}}))) +
      geom_col(fill = color, width = .7) +
      labs(x = NULL, y = NULL, title = title) +
      scale_x_continuous(expand = expansion(mult = c(0, .08))),
    tooltip = c("x", "y")
  ) |> config(displayModeBar = FALSE)
}

# Leadership-level derived measures
theme_metadata <- tribble(
  ~theme, ~type, ~impact, ~decision,
  "Clear and useful", "Strength", 2, "Retain the clear structure and practical assessment flow.",
  "Scoring needs explanation", "Concern", 5, "Clarify scoring criteria and add worked examples.",
  "Too long or repetitive", "Concern", 3, "Combine overlapping items and offer a shorter screening version.",
  "Technical language", "Concern", 4, "Add plain-language definitions and a glossary.",
  "Needs recommendations", "Concern", 5, "Connect low scores with recommended next steps.",
  "Navigation and saving", "Concern", 3, "Add progress, save confirmation, and question links.",
  "Collaboration and roles", "Strength", 3, "Retain team use and clarify responsibility by section.",
  "Result validity", "Concern", 5, "Explain calculations and show completeness or confidence."
)

theme_leadership <- theme_hits |>
  distinct(respondent_id, theme) |>
  count(theme, name = "respondents") |>
  mutate(percentage = respondents / n_distinct(survey$respondent_id)) |>
  right_join(theme_metadata, by = "theme") |>
  mutate(respondents = replace_na(respondents, 0L), percentage = replace_na(percentage, 0))

top_strengths <- theme_leadership |> filter(type == "Strength") |> arrange(desc(percentage)) |> slice_head(n = 3)
top_concerns <- theme_leadership |> filter(type == "Concern") |> arrange(desc(percentage), desc(impact)) |> slice_head(n = 3)

priority_data <- theme_leadership |>
  filter(type == "Concern") |>
  mutate(priority = case_when(
    impact >= 4 & percentage >= median(percentage) ~ "Immediate",
    impact >= 4 ~ "Strategic",
    percentage >= median(percentage) ~ "Quick win",
    TRUE ~ "Monitor"
  ))

question_sentiment <- sentiment_complete |>
  count(question, sentiment_group) |>
  group_by(question) |>
  mutate(percentage = n / sum(n)) |>
  ungroup()

actionability <- survey_long |>
  filter(question == "results") |>
  mutate(category = case_when(
    str_detect(str_to_lower(response), "next step|recommend|action|remediation|did not provide|not provide enough") ~ "Understood, but needs next steps",
    str_detect(str_to_lower(response), "unclear|confus|difficult|not confident|uncertain") ~ "Difficult to interpret",
    str_detect(str_to_lower(response), "useful|informative|clear|easy to see|worked well") ~ "Understandable and useful",
    TRUE ~ "No clear actionability signal"
  )) |>
  count(category) |>
  mutate(percentage = n / sum(n))

readiness_indicators <- tribble(
  ~indicator, ~pattern,
  "Perceived usefulness", "useful|valuable|helpful|practical",
  "Ease of use", "clear|easy|straightforward|simple to use",
  "Result confidence", "confident|accurate|valid|reflected",
  "Actionability", "next step|recommend|action|remediation",
  "Team usefulness", "team|collaborat|discussion|data owner|project manager"
) |>
  rowwise() |>
  mutate(
    respondents = n_distinct(survey_long$respondent_id[str_detect(str_to_lower(survey_long$response), pattern)]),
    percentage = respondents / n_distinct(survey$respondent_id)
  ) |>
  ungroup() |>
  select(-pattern) |>
  bind_rows(tibble(indicator = "Intention to reuse/recommend", respondents = NA_integer_, percentage = NA_real_))

polarization <- sentiment_complete |>
  count(department, sentiment_group) |>
  group_by(department) |>
  mutate(percentage = n / sum(n)) |>
  ungroup() |>
  select(department, sentiment_group, percentage) |>
  pivot_wider(names_from = sentiment_group, values_from = percentage, values_fill = 0) |>
  mutate(
    spread = abs(Positive - Negative),
    interpretation = case_when(
      Positive >= .60 ~ "Broadly positive",
      Negative >= .60 ~ "Broadly negative",
      spread <= .15 ~ "Polarized or mixed",
      Positive > Negative ~ "Leans positive",
      TRUE ~ "Leans negative"
    )
  )

representative_quotes <- theme_hits |>
  semi_join(bind_rows(top_strengths, top_concerns), by = "theme") |>
  group_by(theme) |>
  slice_max(order_by = str_length(response), n = 1, with_ties = FALSE) |>
  ungroup() |>
  transmute(theme, department, quotation = paste0("“", response, "”"))

decision_table <- priority_data |>
  arrange(desc(priority == "Immediate"), desc(impact), desc(percentage)) |>
  transmute(
    Priority = priority,
    Finding = theme,
    Evidence = paste0(percent(percentage, accuracy = 1), " of respondents mentioned this theme"),
    `Recommended decision` = decision,
    Timing = case_when(priority == "Immediate" ~ "Immediate", priority == "Strategic" ~ "Next release", TRUE ~ "Future enhancement")
  )

sentiment_stats <- sentiment_complete |>
  count(sentiment_group) |>
  mutate(percentage = n / sum(n))
safe_sentiment <- function(label) {
  value <- sentiment_stats |> filter(sentiment_group == label) |> pull(percentage)
  if (length(value) == 0) 0 else value[[1]]
}

positive_pct <- safe_sentiment("Positive")
negative_pct <- safe_sentiment("Negative")
top_strength_text <- paste(top_strengths$theme, collapse = ", ")
top_concern_text <- paste(top_concerns$theme, collapse = ", ")
action_text <- paste(top_concerns$decision, collapse = " ")

executive_summary <- c(
  paste0(
    "Overall, ", percent(positive_pct, accuracy = 1),
    " of analyzed comments contained more positive than negative language, while ",
    percent(negative_pct, accuracy = 1),
    " contained more negative language. These percentages are directional indicators because one comment can contain both praise and criticism."
  ),
  paste0(
    "The strongest recurring aspects were ", top_strength_text,
    ". The leading concerns were ", top_concern_text,
    ". This indicates that the scorecard has a useful foundation, but clarity, usability, and confidence in the results may affect consistent adoption."
  ),
  paste0(
    "Leadership should focus first on the high-frequency, high-impact concerns and review whether experiences differ materially across departments. ",
    action_text
  ),
  paste0(
    "The findings are based on transparent keyword and sentiment rules rather than a generative model. They should be confirmed through a small review of representative comments before final decisions are made."
  )
)
```

# Overview

## Row {height=18%}

```{r}
htmltools::div(
  class = "kpi-grid",
  htmltools::div(class = "kpi-card kpi-teal", htmltools::span(class = "kpi-label", "Respondents"), htmltools::strong(n_distinct(survey$respondent_id)), htmltools::small("People represented")),
  htmltools::div(class = "kpi-card kpi-navy", htmltools::span(class = "kpi-label", "Departments"), htmltools::strong(n_distinct(survey$department)), htmltools::small("Organizational perspectives")),
  htmltools::div(class = "kpi-card kpi-blue", htmltools::span(class = "kpi-label", "Open-text responses"), htmltools::strong(comma(nrow(survey_long))), htmltools::small("Comments analyzed")),
  htmltools::div(class = "kpi-card kpi-gold", htmltools::span(class = "kpi-label", "Completion"), htmltools::strong(percent(completion_rate, accuracy = 1)), htmltools::small("Questions answered"))
)
```

## Row {height=42%}

### Executive summary

```{r}
htmltools::div(
  class = "executive-summary executive-summary-compact",
  lapply(executive_summary[1:3], function(paragraph) htmltools::p(paragraph))
)
```

### Leading strengths and concerns

```{r}
leadership_finding <- function(label, value, kind) {
  htmltools::div(
    class = paste("finding-item", kind),
    htmltools::span(class = "finding-label", label),
    htmltools::strong(percent(value, accuracy = 1))
  )
}

htmltools::div(
  class = "findings-panel",
  htmltools::h4("Strengths"),
  lapply(seq_len(nrow(top_strengths)), function(i) leadership_finding(top_strengths$theme[i], top_strengths$percentage[i], "strength")),
  htmltools::h4("Concerns"),
  lapply(seq_len(nrow(top_concerns)), function(i) leadership_finding(top_concerns$theme[i], top_concerns$percentage[i], "concern"))
)
```

## Row {height=40%}

### Recommended decisions

```{r}
overview_decisions <- decision_table |> slice_head(n = 4)
htmltools::div(
  class = "decision-list",
  lapply(seq_len(nrow(overview_decisions)), function(i) {
    htmltools::div(
      class = "decision-item",
      htmltools::span(class = paste("priority-badge", tolower(overview_decisions$Priority[i])), overview_decisions$Priority[i]),
      htmltools::div(htmltools::strong(overview_decisions$Finding[i]), htmltools::p(overview_decisions$`Recommended decision`[i]))
    )
  })
)
```

### Representative comments

```{r}
overview_quotes <- representative_quotes |> slice_head(n = 3)
htmltools::div(
  class = "quote-list",
  lapply(seq_len(nrow(overview_quotes)), function(i) {
    htmltools::blockquote(
      overview_quotes$quotation[i],
      htmltools::footer(paste(overview_quotes$theme[i], "•", overview_quotes$department[i]))
    )
  })
)
```

# Priorities

## Row

### 1. Executive sentiment summary

```{r}
sentiment_executive <- sentiment_stats |>
  mutate(sentiment_group = factor(sentiment_group, levels = c("Positive", "Mixed/neutral", "Negative")))

ggplotly(
  ggplot(sentiment_executive, aes(sentiment_group, percentage, fill = sentiment_group,
                                  text = paste0(percent(percentage, accuracy = 1), " of comments"))) +
    geom_col(width = .65) +
    scale_fill_manual(values = c("Positive" = teal, "Mixed/neutral" = amber, "Negative" = red)) +
    scale_y_continuous(labels = percent) +
    labs(x = NULL, y = NULL, title = "Overall experience indicator") +
    theme(legend.position = "none"),
  tooltip = "text"
) |> config(displayModeBar = FALSE)
```

### 2. Top strengths and concerns

```{r}
strength_concern_table <- bind_rows(top_strengths, top_concerns) |>
  transmute(Type = type, Finding = theme, Respondents = respondents, Percentage = percentage)

datatable(
  strength_concern_table,
  rownames = FALSE,
  options = list(dom = "t", autoWidth = TRUE)
) |>
  formatPercentage("Percentage", 0)
```

## Row

### 3. Leadership priority matrix

```{r}
ggplotly(
  ggplot(priority_data, aes(percentage, impact, color = priority,
                            text = paste0(theme, "<br>Frequency: ", percent(percentage, accuracy = 1),
                                          "<br>Impact: ", impact, "/5<br>Priority: ", priority))) +
    geom_point(size = 5, alpha = .9) +
    geom_text(aes(label = theme), nudge_y = .12, size = 3, show.legend = FALSE) +
    scale_x_continuous(labels = percent, limits = c(0, max(priority_data$percentage) * 1.15)) +
    scale_y_continuous(breaks = 1:5, limits = c(1, 5.4)) +
    scale_color_manual(values = c("Immediate" = red, "Strategic" = amber, "Quick win" = teal, "Monitor" = gray)) +
    labs(x = "Percentage of respondents", y = "Business impact", color = "Priority",
         title = "Frequency and impact of concerns"),
  tooltip = "text"
) |> config(displayModeBar = FALSE)
```

### 4. Department comparison

```{r}
department_leadership <- theme_hits |>
  filter(theme %in% top_concerns$theme) |>
  distinct(department, respondent_id, theme) |>
  count(department, theme) |>
  group_by(department) |>
  mutate(percentage = n / n_distinct(survey$respondent_id[survey$department == first(department)])) |>
  ungroup()

ggplotly(
  ggplot(department_leadership, aes(department, theme, fill = percentage,
                                    text = paste0(percent(percentage, accuracy = 1), " of department respondents"))) +
    geom_tile(color = "white", linewidth = 1) +
    scale_fill_gradient(low = "#EFF6FF", high = navy, labels = percent) +
    labs(x = NULL, y = NULL, fill = "Percentage", title = "Leading concerns by department") +
    theme(axis.text.x = element_text(angle = 30, hjust = 1)),
  tooltip = "text"
) |> config(displayModeBar = FALSE)
```

# Performance

## Row

### 5. Question-level performance

```{r}
ggplotly(
  ggplot(question_sentiment, aes(question, percentage, fill = sentiment_group,
                                 text = paste0(sentiment_group, ": ", percent(percentage, accuracy = 1)))) +
    geom_col(position = "fill") +
    scale_fill_manual(values = c("Positive" = teal, "Mixed/neutral" = amber, "Negative" = red)) +
    scale_y_continuous(labels = percent) +
    labs(x = "Question", y = NULL, fill = "Sentiment", title = "Sentiment by scorecard component"),
  tooltip = "text"
) |> config(displayModeBar = FALSE)
```

### 6. Actionability of the output

```{r}
plotly_bar(actionability, percentage, category, "Can users act on the scorecard results?", teal) |>
  layout(xaxis = list(tickformat = ".0%"))
```

## Row

### 7. Adoption and readiness indicators

```{r}
readiness_display <- readiness_indicators |>
  mutate(
    Result = if_else(is.na(percentage), "Not directly measured", percent(percentage, accuracy = 1)),
    Note = if_else(is.na(percentage), "Add a direct survey question in the next round", "Text-based proxy")
  ) |>
  select(Indicator = indicator, Result, Note)

datatable(readiness_display, rownames = FALSE, options = list(dom = "t", autoWidth = TRUE))
```

### 8. Agreement versus polarization

```{r}
polarization_display <- polarization |>
  transmute(
    Department = department,
    Positive = Positive,
    Negative = Negative,
    Interpretation = interpretation
  )

datatable(polarization_display, rownames = FALSE, options = list(dom = "t", autoWidth = TRUE)) |>
  formatPercentage(c("Positive", "Negative"), 0)
```

# Decisions

## Row

### 9. Representative comments

```{r}
datatable(
  representative_quotes,
  rownames = FALSE,
  escape = TRUE,
  options = list(pageLength = 8, dom = "tip", autoWidth = TRUE),
  colnames = c("Theme", "Department", "Anonymized quotation")
)
```

### 10. Recommended decisions

```{r}
datatable(
  decision_table,
  rownames = FALSE,
  options = list(pageLength = 8, dom = "tip", autoWidth = TRUE)
)
```

# Themes

## Row

### Themes by survey question

```{r}
theme_question <- theme_hits |>
  count(question, theme) |>
  mutate(question = factor(question, levels = names(question_map)))

ggplotly(
  ggplot(theme_question, aes(question, theme, fill = n, text = paste("Mentions:", n))) +
    geom_tile(color = "white", linewidth = 1) +
    scale_fill_gradient(low = "#E6FFFB", high = teal) +
    labs(x = "Question", y = NULL, fill = "Mentions", title = "Theme heat map"),
  tooltip = "text"
) |> config(displayModeBar = FALSE)
```

### Themes by department

```{r}
theme_department <- theme_hits |> count(department, theme)
ggplotly(
  ggplot(theme_department, aes(department, theme, fill = n, text = paste("Mentions:", n))) +
    geom_tile(color = "white", linewidth = 1) +
    scale_fill_gradient(low = "#EFF6FF", high = navy) +
    labs(x = NULL, y = NULL, fill = "Mentions", title = "Department comparison") +
    theme(axis.text.x = element_text(angle = 30, hjust = 1)),
  tooltip = "text"
) |> config(displayModeBar = FALSE)
```

# Language

## Row

### Frequent words

```{r}
word_summary <- words |> count(word, sort = TRUE) |> slice_head(n = 20)
plotly_bar(word_summary, n, word, "Top words", navy)
```

### Frequent phrases

```{r}
bigram_summary <- bigrams |> group_by(question) |> slice_max(n, n = 5, with_ties = FALSE) |> ungroup()
datatable(
  bigram_summary,
  rownames = FALSE,
  filter = "top",
  options = list(pageLength = 12, dom = "tip", autoWidth = TRUE),
  colnames = c("Question", "Two-word phrase", "Count")
)
```

# Comments

## Row

### Searchable response table

```{r}
comment_table <- survey_long |>
  mutate(question = as.character(question)) |>
  select(respondent_id, department, question, response)

datatable(
  comment_table,
  rownames = FALSE,
  filter = "top",
  escape = TRUE,
  options = list(pageLength = 15, scrollX = TRUE, autoWidth = TRUE),
  colnames = c("Respondent", "Department", "Question", "Response")
)
```
